---
title: "R Notebook"
output: html_notebook
---


```{r}
library(tidyverse)
require(foreign)
require(cluster)
require(factoextra)
require(cluster)
require(mclust)
require(lcmm)
library(flexmix)
```

WITH THE LONG DATA SET

```{r}
table_3 <- read_fst("C:/Users/aslop/Desktop/SPRING 2019/Supervised and Unsupervised Machine Learning/FINAL Project/Papers for Research/Final data sets/table_3_long.fst")


head(table_3)
```

```{r}
data_long <- table_3 %>% 
  mutate(time = case_when(
    year == 2010 ~ 1,
    year == 2011 ~ 2,
    year == 2012 ~ 3,
    year == 2013 ~ 4,
    year == 2014 ~ 5,
    year == 2015 ~ 6,
    year == 2016 ~ 7,
    year == 2017 ~ 8
  )) %>% 
  select(year, time, puma, diversion_rate_total, below_pover, nodipl_rate, everything())


head(data_long)
glimpse(data_long)
```


```{r}
data <- data_long %>% 
  select(puma, boro, borough, diversion_rate_total, nodipl_rate, below_pover, year) %>% 
  filter(year == 2017)


glimpse(data)
```



```{r}
fit.nagin.re_data <- fit.nagin.re$pprob

fit.nagin.re_data

```






```{r}
fit.nagin.re <- hlme(diversion_rate_total~time+I(time^2),
                     mixture=~time+I(time^2), 
                     classmb=~1, random=~1,subject='puma',ng=3, 
                     data=data_long)

head(fit.nagin.re, 20)
```


```{r}
plot(fit.nagin.re,which='fit',var.time='time',legend=NULL)
```

'



MODEL 1 -- linear model, two clustes
###################


```{r}
### latent class linear mixed model with 2 classes

mod_1<-hlme(diversion_rate_total~time, mixture=~time,random=~time, subject='puma',
         ng=2,data=data_long)

summary(mod_1)
```


```{r}
plot(mod_1,which='fit',var.time='time',legend=NULL)
```


MODEL 2 -- linear model, three clustes


```{r}
mod_2<-hlme(diversion_rate_total~time, mixture=~time,random=~time, subject='puma',
         ng=3,data=data_long)

summary(mod_2)


```

```{r}
plot(mod_2,which='fit',var.time='time',legend=NULL)
```

```{r}

```



MODEL 3 -- linear model, four clustes


```{r}
mod_3<-hlme(diversion_rate_total~time, mixture=~time,random=~time, subject='puma',
         ng=4,data=data_long)

summary(mod_3)


```

```{r}
plot(mod_3,which='fit',var.time='time',legend=NULL)
```


MODEL  --- three clusters quadratic term

```{r}
mod_4 <- hlme(diversion_rate_total~time+I(time^2),
                     mixture=~time+I(time^2), 
                     classmb=~1, random=~1,subject='puma',ng=3, 
                     data=data_long)

summary(mod_4)
```



```{r}
plot(mod_4, which='fit',var.time='time',legend=NULL)
```



MODEL 5 --- three clusters cubic term

```{r}
mod_5 <- hlme(diversion_rate_total~time+I(time^3),
                     mixture=~time+I(time^3), 
                     classmb=~1, random=~1,subject='puma',ng=3, 
                     data=data_long)
summary(mod_5)

```


```{r}
plot(mod_5, which='fit',var.time='time',legend=NULL)
```


PROFILE OF THE LINEAR HLME MODEL WITH THREE CLUSTERS

```{r}
model_2_table <- mod_2$pprob

data_mod2 <- model_2_table %>% 
  left_join(data, by = "puma")


table_mod2 <- data_mod2 %>% 
  group_by(class) %>% 
  summarise(diver_rate = mean(diversion_rate_total),
            nodipl_rate = mean(nodipl_rate),
            poverty = mean(below_pover)) %>% 
  arrange(desc(diver_rate))


table_mod2
```


```{r}
table_mod2 <- data_mod2 %>% 
  group_by(class, borough) %>% 
  summarise(count = length(borough))

table_mod2 
```










```{r}
model_4_table <- mod_4$pprob

head(model_4_table)

glimpse(model_4_table)
```

Clusters of the tree cluster quadratic model with covariates for 2017

```{r}
data_table <- model_4_table %>% 
  left_join(data, by = "puma")

glimpse(data_table)
```


Profile of the model 4 ( three cluster, quadratic model)

```{r}
table_1 <- data_table %>% 
  group_by(class) %>% 
  summarise(diver_rate = mean(diversion_rate_total),
            nodipl_rate = mean(nodipl_rate),
            poverty = mean(below_pover)) %>% 
  arrange(desc(diver_rate))


table_1
```


```{r}
table_2 <- data_table %>% 
  group_by(class, borough) %>% 
  summarise(count = length(borough))

table_2 
  
```

TRAJ WITH THREE CLUSTERS AND QUADRATIC TERMS AND NO DIPLOME MEASUREMENT

```{r}
traj_nodipl <- read_csv("C:/Users/aslop/Desktop/SPRING 2019/Supervised and Unsupervised Machine Learning/FINAL Project/Papers for Research/Final data sets/traj_nodipl.csv")
```

```{r}
data_nodipl <- traj_nodipl %>% 
  select(puma, `_traj_Group`)

data_nodipl <- data_nodipl %>% 
  left_join(data, by = "puma")
```



```{r}
table_nodipl <- data_nodipl %>% 
  group_by(`_traj_Group`) %>% 
  summarise(diver_rate = mean(diversion_rate_total),
            nodipl_rate = mean(nodipl_rate),
            poverty = mean(below_pover)) %>% 
  arrange(desc(diver_rate))


table_nodipl
```


```{r}
table_nodipl_2 <- data_nodipl %>% 
  group_by(`_traj_Group`, borough) %>% 
  summarise(count = length(borough))

table_nodipl_2
```

>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>


```{r}
traj_plot <- traj_nodipl %>% 
  select(puma, `_traj_Group`) %>% 
  left_join(table_3, key = "puma")
  
```


```{r}
# plot the frame of axes

plot_1 <- traj_plot %>% 
  filter(`_traj_Group` == 1)


plot_2 <- traj_plot %>% 
  filter(`_traj_Group` == 2)


plot_3 <- traj_plot %>% 
  filter(`_traj_Group` == 3)





plot (0, 0, type="n", xlim=c(2010, 2017), ylim=c(5, 60), main="NYC Sub-areas waste diversion trajectories 2010 - 2017", 
     ylab="Diversion Rate", xlab="", las=1, lwd=2, bty="n", cex.axis=0.7)

codes <- unique(plot_1$puma)
for (i in 1:length(codes)) {
    currCountry <- subset(plot_1, puma == codes[i])
    currCountry <- currCountry[order(currCountry$year, decreasing=FALSE),]
    lines(2010:2017, currCountry$diversion_rate_total, col="green")
}

codes <- unique(plot_2$puma)
for (i in 1:length(codes)) {
    currCountry <- subset(plot_2, puma == codes[i])
    currCountry <- currCountry[order(currCountry$year, decreasing=FALSE),]
    lines(2010:2017, currCountry$diversion_rate_total, col="red")
}


codes <- unique(plot_3$puma)
for (i in 1:length(codes)) {
    currCountry <- subset(plot_3, puma == codes[i])
    currCountry <- currCountry[order(currCountry$year, decreasing=FALSE),]
    lines(2010:2017, currCountry$diversion_rate_total, col="blue")
}
```








Comparing to the three model quadratic model with traj


```{r}
traj_quad_stata <- read_csv("C:/Users/aslop/Desktop/SPRING 2019/Supervised and Unsupervised Machine Learning/FINAL Project/Papers for Research/Final data sets/traj_quad_stata.csv")
```

```{r}
data_traj <- traj_quad_stata %>% 
  select(puma, `_traj_Group`)

data_traj <- data_traj %>% 
  left_join(data, by = "puma")

```


```{r}
table_traj <- data_traj %>% 
  group_by(`_traj_Group`) %>% 
  summarise(diver_rate = mean(diversion_rate_total),
            nodipl_rate = mean(nodipl_rate),
            poverty = mean(below_pover)) %>% 
  arrange(desc(diver_rate))


table_traj
```



```{r}
table_2_traj <- data_traj %>% 
  group_by(`_traj_Group`, borough) %>% 
  summarise(count = length(borough))

table_2_traj
```


```{r}

```










HLME
- heterogeneous populations (relax (iv))
→ hlme for latent class linear mixed models (i.e. Gaussian continuous
outcome)






MODEL with time dependent variables - quadratic term + no diploma covariate

```{r}
mod_6 <- hlme(diversion_rate_total~time+I(time^2),
                     mixture=~time+I(time^2), 
                     classmb=~nodipl_rate, random=~1,subject='puma',ng=3, 
                     data=data_long)

summary(mod_6)

```



```{r}
plot(mod_6, which='fit',var.time='time',legend=NULL)
```



MODEL with time dependent variables - quadratic term + no diploma covariate

```{r}
mod_7 <- hlme(diversion_rate_total~time+I(time^3),
                     mixture=~time+I(time^3), 
                     classmb=~nodipl_rate, random=~1,subject='puma',ng=3, 
                     data=data_long)

summary(mod_7)
```

```{r}
plot(mod_7, which='fit',var.time='time',legend=NULL)
```






```{r}
mod_8 <- hlme(diversion_rate_total~time+I(time^2),
                     mixture=~time+I(time^2), 
                     classmb=~below_pover, random=~1,subject='puma',ng=3, 
                     data=data_long)

summary(mod_8)
```

```{r}
plot(mod_8, which='fit',var.time='time',legend=NULL)
```




















USING FLEXMIX

```{r}
p <- flexmix(diversion_rate_total ~ time | puma, k=2,
             model=FLXMRglm(family="gaussian"),
             data=data_long)

summary(p)
```

```{r}
prior(p)
```

```{r}
parameters(p)
```


```{r}
 # Extract each component
comp_1 <- parameters(p, component=1)
comp_2 <- parameters(p, component=2)
```


```{r}
# Extract the means
mean_comp_1 <- comp_1[1:2]
mean_comp_2 <- comp_2[1:2]
```



```{r}
 # Extract the covariance matrices
covariance_comp_1 <- matrix(comp_1[3:3], nrow=2)
covariance_comp_2 <- matrix(comp_2[3:6], nrow=2)
covariance_comp_1
```

```{r}
clusters(p)
```


Bivariate distribution

```{r}
p2 <- flexmix(cbind(diversion_rate_total, below_pover) ~ time | puma, k=2,
             model=FLXMCmvnorm(diag = TRUE),
             data=data_long)

# gives the proportions of each cluster
prior(p2)
```

```{r}
# two first rows - means and the covariance matrices formed by the following 4 values
parameters((p2))
```

